Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Validation

### 2012-11-06 paper.tex
Added test subsections
Filled out gravitational free fall section, added directory with results for this simulation

%==========================================================================
% Title
%==========================================================================
\title{WD Merger Notes}

\shorttitle{WD Mergers. I. Methodology}
\shortauthors{Max}

@@

>>>>

%==========================================================================
% Title
%==========================================================================
\title{Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Validation}

\shorttitle{DWD Mergers. I. Methodology}
\shortauthors{Max}

@@

\subsection{Maintaining Hydrostatic Equilibrium}\label{Sec:HSE}

\subsection{Gravitational Free Fall}\label{Sec:Gravitational Free Fall}

A simple test to verify the Poisson solver implemented by CASTRO is the case of gravitational free fall. In this setup, two stars, each individually in an equilibrium state, are placed on the computational grid, separated by an initial distance $r_0$ along the $x$ axis with zero initial velocity. We choose stars of masses $0.6$ and $0.8\, M_\odot$, with the lower mass star on the left side (e.g. $x < 0$) of the domain and the higher mass star on the right, such that their center of mass coincides with the center of the domain. Gravitationally, the stars may be treated as point masses until the point of contact, so the equation of motion governing the distance $r$ between their centers of mass is that of simple free fall:
\begin{equation}
  \ddot{r}(t) = - \frac{GM}{r},
\end{equation}
where $G$ is the gravitational constant and $M$ is the total mass of the system. This differential equation has a closed-form solution for the evolution time as a function of separation:
\begin{equation}
  t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
\end{equation}
We determine the initial separation to be consistent with the simulation performed in Section \ref{Sec:Kepler}; that is, we select an initial orbital period and use Kepler's third law to calculate the radius of a circular orbit with that period. We select an initial orbital period of 200 s for this simulation, so that the separation is
\[
  r_0 = 5.73 \times 10^{9}\ \text{cm}.
\]
We chose a relatively low resolution simulation to demonstrate the capabilities of CASTRO even while using modest resources. The computational grid is covered by a coarse grid of $48^3$ zones, with two levels of refinement above the coarse grid. Each refined grid carries an increase in resolution by a factor of 4 relative to the coarser grid below it. CASTRO initially assigns $93\%$ of the domain to be covered by the intermediate resolution grids, and $0.05\%$ of the domain to be covered by the finest resolution grids.

The analytical result in Equation \ref{analyticalFreeFall} determines the total elapsed free-fall time,
\[
  t_{\text{ff}} = \frac{\pi}{2} \sqrt{\frac{r_0^3}{2GM}}.
\]
The physical radius of each white dwarf is roughly $10\%$ of the initial separation, so we consider the evolution terminated when the radial separation reaches that value (at that point, the stars will no longer be in free-fall due to physical contact). Since the elapsed time goes roughly as the square root of the distance for small $r$, we consider the evolution terminated when $t = 0.99\, t_{\text{ff}}$. The results of our simulation are shown in Figure \ref{Fig:Free Fall}.

\begin{figure}
  \centering
  \includegraphics[scale=2.0]{freefall/plot_freefall}
  \caption{Time evolution of two initially stationary white dwarfs, mutually attracted to each other by the gravitational force. The horizontal axis gives the separation of the white dwarfs, scaled to the initial separation, and the vertical axis gives the elapsed time of the simulation, scaled to the total elapsed time necessarily. The solid curve shows the analytical result, calculated from Newtonian mechanics, and the circles show the samples from the time evolution with CASTRO. The agreement between theory and simulation is excellent, even for a simulation with modest resolution.}
  \label{Fig:Free Fall}
\end{figure}

\subsection{Keplerian Orbit}\label{Sec:Kepler}
